Greenhouse gas emissions model

### 2021-10-04 210617_GHG_model.Rmd
Add updates from the uncertainty package (mostly just labels for the plots)

```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1}
mermaid("
        graph LR
        C(Central collection)-->B(Burned); linkStyle 0 stroke: red, stroke-width:1px
                                B-->P(Pasture); linkStyle 1 stroke: red, stroke-width:1px
        C-->D(Daily spread); linkStyle 2 stroke: green, stroke-width:1px
            D-->S(Sold); linkStyle 3 stroke: red, stroke-width:1px
                S-->L(Liquid storage); linkStyle 4 stroke: red, stroke-width:1px
                    L-->Co(Composted); linkStyle 5 stroke: red, stroke-width:1px
                        Co-->Dl(Dry lot); linkStyle 6 stroke: red, stroke-width:1px
                             Dl-->St(Solid storage); linkStyle 7 stroke: red, stroke-width:1px
                    L-->Bi(Biogas); linkStyle 8 stroke: green, stroke-width:1px
                        Bi-->Ls(Liquid slurry); linkStyle 9 stroke: red, stroke-width:1px
        ")
```

@@

```{r plot_distributions, fig.width=10, fig.height=10}
# enteric_CH4=enteric_CH4,
enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Total Methane Emissions from Livestock by Enteric Fermentation") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

@@

Below we apply a post-hoc analysis to the `mcSimulation()` outputs with the `plsr.mcSimulation()` function in the `decisoonSupport` package. We use the function to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model. This function uses the outputs of the `mcSimulation()` selecting all the input variables from the `GHG_simulation` decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`. We use the code `names(GHG_simulation$y)[7]` to select the outcome variable `r names(GHG_simulation$y)[7]`, which is an element of the list `y` in our `GHG_simulation` outputs.

@@

We apply a collection of functions for use in assessing uncertainty in the relationship between GHG emissions and milk yields in livestock production. We use functions from the `uncertainty` package. Install the package using `devtools::install_github("CWWhitney/uncertainty")`.

We apply the uncertainty functions on the `milk_yield` and `on_farm` variables. These come from the model results (`r nrow(GHG_simulation$y)`) of the `GHG_simulation` model results to the `in_var` (x) and an outcome variable `out_var` (y). in the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).

```{r name_in_out_vars}
in_var  <- GHG_simulation$y$milk_yield
out_var <- GHG_simulation$y$on_farm
```

@@

Here we use the `varscatter()` function to create a scatter plot of milk yields `in_var` (x) and GHG emissions `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves. Plot made with the scatter.hist() function in the `plyr()` package [@R-plyr] in R [@R-base]. The result is a scatter plot with associated and estimated densities (using a loess smooth) given by the red dot (mean) and red ellipses (1 and 2 sigma from mean). The red line going across the plot shows the linear fit. Histograms are shown with smooth lines (loess smooth linear fits) density curves. The numeric value in the upper right gives the Spearman correlation coefficient between the influencing variable and the outcome variable.

```{r uncertainty_package}
uncertainty::varscatter(in_var = in_var,
                        out_var = out_var)
```

@@

```{r varkernel}
uncertainty::varkernel(in_var = in_var, out_var = out_var)
```

@@

We use the `varkernelslicerange()` function to show the milk yield (influencing variable `in_var` value) intervals. Here we choose a very wide distribution (the whole range of milk yields from the Monte Carlo simulation). Here we set the maximum expected value for milk yield `in_var` to `r round(max(in_var), digits = 0)` and the minimum expected value to `r round(min(in_var), digits = 0)`.

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = round(min(in_var)),
                                 max_in_var = round(max(in_var)))
```

@@

For a more specific estimate we can choose from those optimized interquartile ranges to select a range to slice from the density kernel `varkernel()`. we do that here using the `varviolin()` function to determine meaningful milk yield variable intervals and how they interact with the outcome variable GHG emissions. We do this by calculating an 'optimal' interval width using the interquartile range `IQR()` function in the `stats()` package.

```{r varviolin, fig.width=17, fig.height=7}
uncertainty::varviolin(in_var = in_var, out_var = out_var, legend.position = "left")
```

>>>>

```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1, include = F}
mermaid("
        graph LR
        C(Central collection)-->B(Burned); linkStyle 0 stroke: red, stroke-width:1px
                                B-->P(Pasture); linkStyle 1 stroke: red, stroke-width:1px
        C-->D(Daily spread); linkStyle 2 stroke: green, stroke-width:1px
            D-->S(Sold); linkStyle 3 stroke: red, stroke-width:1px
                S-->L(Liquid storage); linkStyle 4 stroke: red, stroke-width:1px
                    L-->Co(Composted); linkStyle 5 stroke: red, stroke-width:1px
                        Co-->Dl(Dry lot); linkStyle 6 stroke: red, stroke-width:1px
                             Dl-->St(Solid storage); linkStyle 7 stroke: red, stroke-width:1px
                    L-->Bi(Biogas); linkStyle 8 stroke: green, stroke-width:1px
                        Bi-->Ls(Liquid slurry); linkStyle 9 stroke: red, stroke-width:1px
        ")
```

@@

```{r plot_distributions, fig.width=10, fig.height=10, include = F}
# add , include = F to keep this distribution figure building step out of the RMD
# enteric_CH4=enteric_CH4,
enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Total Methane Emissions from Livestock by Enteric Fermentation") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

@@

Below we apply a post-hoc analysis to the `mcSimulation()` outputs with the `plsr.mcSimulation()` function in the `decisionSupport` package. We use the function to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model. This function uses the outputs of the `mcSimulation()` selecting all the input variables from the `GHG_simulation` decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`. We use the code `names(GHG_simulation$y)[7]` to select the outcome variable `r names(GHG_simulation$y)[7]`, which is an element of the list `y` in our `GHG_simulation` outputs.

@@

Here we apply a collection of functions for use in assessing uncertainty in the relationship between GHG emissions and milk yields in livestock production. We use functions from the `uncertainty` package. Install the package using `devtools::install_github("CWWhitney/uncertainty")`.

We apply the uncertainty functions on the `milk_yield` and `on_farm` variables. These come from the model results (`r nrow(GHG_simulation$y)`) of the `GHG_simulation` model results to the `in_var` (x) and an outcome variable `out_var` (y). In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).

```{r name_in_out_vars}
in_var  <- GHG_simulation$y$milk_yield # "Total Milk Yield"
out_var <- GHG_simulation$y$on_farm # "Total emissions"
```

@@

Here we use the `varscatter()` function to create a scatter plot of milk yields `in_var` (x) and GHG emissions `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves. We created the plot with the `scatter.hist()` function in the `plyr()` package [@R-plyr] in R [@R-base]. The result is a scatter plot with associated and estimated densities (using a loess smooth) given by the red dot (mean) and red ellipses (1 and 2 sigma from mean). The red line going across the plot shows the linear fit. Histograms are shown with smooth lines (loess smooth linear fits) density curves. The numeric value in the upper right gives the Spearman correlation coefficient between the influencing variable and the outcome variable.

```{r uncertainty_package}
uncertainty::varscatter(in_var = in_var,
                        out_var = out_var,
                        xlab = "Total Milk Yield",
                        ylab = "Total emissions")
```

@@

```{r varkernel}
uncertainty::varkernel(in_var = in_var, 
                       out_var = out_var,
                        xlab = "Total Milk Yield",
                        ylab = "Total emissions")
```

@@

We use the `varkernelslicerange()` function to show the total emissions given a specified milk yield (influencing variable `in_var` value) interval. Here we choose a very wide distribution (the whole range of milk yields from the Monte Carlo simulation). Here we set the maximum expected value for milk yield `in_var` to the maximum (`r round(max(in_var), digits = 0)`) and the minimum expected value (`r round(min(in_var), digits = 0)`) from our simulation. We can also choose another range by supplying milk yield values to `max_in_var` and `min_in_var`.

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, 
                                 out_var = out_var, 
                                 min_in_var = round(min(in_var)),
                                 max_in_var = round(max(in_var)), 
                                 xlab_vars = "Total emissions given milk yield range")
```

@@

For a more specific estimate we can choose from those optimized interquartile ranges to select a range to slice from the density kernel `varkernel()`. We do that here using the `varviolin()` function to determine meaningful milk yield variable intervals and how they interact with the outcome variable GHG emissions. We do this by calculating an 'optimal' interval width using the interquartile range `IQR()` function in the `stats()` package.

```{r varviolin, fig.width=17, fig.height=7}
uncertainty::varviolin(in_var = in_var, 
                       out_var = out_var, 
                       legend.position = "left",
                        xlab = "Total Milk Yield",
                        ylab = "Total emissions")
```
